Timeliness of security updates for Android

### 2014-02-18 securityupdates.tex
Try and work in argument about lack of information making root exploits problematic

\newcommand{\percMarketShare}{XX\%~\cite{TODO}}
\newcommand{\percInsec}{XX\%~\cite{TODO}}
\newcommand{\daNumManufacturers}{XX~[\S\ref{sec:results}]}
\newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}

@@

@@

\subsection{Root exploits}
We compiled a list of root exploit vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device.
This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in ??? 2013 meant that applications could pretend to be signed with system keys and hence gain root privileges.

We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities.

@@

\subsection{Versions of Android running on devices}
We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
The Device Analyzer data contains data from ??? devices with a total of ??? device days, it contains data for more than ??? months for ??? devices.
%TODO talk about the Device Analyzer project a bit
Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
Among these are the build string and API version for the version of Android currently running on the device each day.
The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.

>>>>

\newcommand{\percMarketShare}{XX\%~\cite{TODO}}
\newcommand{\percInsec}{XX\%~\cite{TODO}}
\newcommand{\daNumManufacturers}{XX~[\S\ref{sec:results}]}
\newcommand{\daNumNetworkOperators}{XXX~[\S\ref{sec:results}]}
\newcommand{\daNumDevices}{XXXX}
\newcommand{\daDeviceDays}{XXXXX}
\newcommand{\daMonths}{X}% Number of months for which we have the below number of devices
\newcommand{\daMonthsDevices}{XXX}
\newcommand{\daAdbPerc}{X\%}

@@

Part of the reason why deployed devices do not receive updates in a timely manner is that there is little data on what is actually happening and so it is not possible to determine which manufacturers or operators are doing a better job.
Hence there is little incentive for manufacturers and operators to do a better job, we will discuss this further in \S\ref{sec:economics}

@@

\subsection{Root exploits}
We compiled a list of root exploit vulnerabilities for Android containing information on when they were known about, what versions they affected and which versions fixed the problem.
We only looked for \emph{root equivalent} exploits which did not require USB debugging to exploit.
\emph{Root equivalent} means that it grants privileges equivalent in scope to root which can then be used to gain root.
Some phones can be `rooted' by enabling USB debugging and using the special privileges of the adb shell to root the device but only \daAdbPerc\ of devices have USB debugging enabled.
This is not something that applications running on the phone can exploit to break compartmentalisation and so we do not include those exploits.
Some exploits which grant root are not traditional kernel exploits, for example the discovery of flaws in the verification of signatures on Android applications in February 2013~\cite{Forristal2013} meant that applications could pretend to be signed with system keys and hence gain root privileges.

We have published full details of the vulnerabilities listed below in an accompanying technical report~\cite{TODO} and summarise them in Table \ref{tab:andvulns}.
We also have a website\footnote{\url{http://androidvulnerabilities.org/}} which we will keep up to date with future vulnerabilities and which has successfully solicited submissions from others.

@@

\subsection{Versions of Android running on devices}
We also need historical information on which versions of Android were running on devices along with other information about those devices which allows us to go into more detail and confirm the reliability of the data our results depend on.
The only suitable data which is available for this is the Device Analyzer data~\cite{Wagner2013}.
The Device Analyzer data contains data from \daNumDevices\ devices with a total of \daDeviceDays\ device days, it contains data for more than \daMonths months for \daMonthsDevices\ devices.
%TODO talk about the Device Analyzer project a bit
Various different kinds of data\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/collected.htm}} are collected by the Device Analyzer Android app\footnote{\url{https://deviceanalyzer.cl.cam.ac.uk/}}.
Among these are the build string and API version for the version of Android currently running on the device each day.
The API/SDK version is a well defined integer, however it does not always change with new Android releases, particularly security bug fixes should not increase this value.
The build string is `The user-visible version string', fortunately most (99.9\%) devices in the data have a build string of the form `x.y.z random string' and so it is possible to extract the Android version number.
